---
title: "tRNA mismatches"
output: html_document
date: "2022-10-25"
---
We want to analyze the tRNA sequencing data in a bit more detail. Let's read it in with this function, and add a sample identification column to each dataframe before binding them together.
```{r}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(cowplot)
library(reshape2)
library(here)

# little read in function, would be better if we'd captured sample IDs in these but OK
read_nanoRMS <- function(x) {
  df <- readr::read_csv(x)
  df %>% 
    select(-std_q, -ACGT, -mean_q)  -> df
}

read_nanoRMS(here("data/basefreq_csvs/CRY1.bwa.sc3.tRNome.per.site.baseFreq.csv")) %>%
  add_column(sample = "wt") -> wt

read_nanoRMS(here("data/basefreq_csvs/w30310x.bwa.sc3.tRNome.per.site.baseFreq.csv")) %>%
  add_column(sample = "bypass") -> bypass

read_nanoRMS(here("data/basefreq_csvs/tpt1d.bwa.sc3.tRNome.per.site.baseFreq.csv")) %>%
  add_column(sample = "tpt1") -> tpt1

read_nanoRMS(here("data/basefreq_csvs/xrn1d10x_untr_tRNA.bwa.sc3.tRNome.per.site.baseFreq.csv")) %>%
  add_column(sample = "xrn1") -> xrn1

read_nanoRMS(here("data/basefreq_csvs/xrn1dtpt1d10x_untr_tRNA.bwa.sc3.tRNome.per.site.baseFreq.csv")) %>%
  add_column(sample = "xrn1tpt1") -> xrn1tpt1

bind_rows(wt, bypass, tpt1, xrn1, xrn1tpt1) -> all

```

Pivot wider & do relevant biological comparisons.
```{r}
all %>%
  select(`#Ref`, pos, strand, mis, sample, cov) %>%
  dplyr::rename(RNA = `#Ref`) %>%
  pivot_wider(names_from = sample, values_from = c(mis, cov)) %>%
  mutate(diff1 = mis_bypass - mis_wt) %>%
  mutate(diff2 = mis_tpt1 - mis_bypass) %>%
  mutate(diff3 = mis_xrn1tpt1 - mis_xrn1) %>%
  group_by(RNA, strand) %>%
  arrange(., pos) %>%
  mutate(mincov1 = pmin(cov_bypass, cov_wt)) %>%
  mutate(mincov2 = pmin(cov_tpt1, cov_bypass)) %>%
  mutate(mincov3 = pmin(cov_xrn1tpt1, cov_xrn1)) %>%
  mutate(adjpos = (pos - 18)) %>% # adjust position numbering so we don't start counting at 5´-adapter sequence
  separate(RNA, c("sac", "cer", "tRNA"), sep = "_") %>%
  separate(tRNA, c("t", "AA", "anticodon", "family", "number")) %>%
  select(-sac, -cer, -t, -number, -family) -> mismatchdiff
```

Calculate minimum coverage requirements: Need to transform select values to NAs before plotting if either mismatch frequency input used for the difference calculation is associated with <20X coverage at that position.

Then we cut the tibble down to remove extraneous information before reshaping & plotting the data.
```{r}
mismatchdiff %>%
  mutate(diff1 = replace(diff1, mincov1 < 20, NA)) %>%
  mutate(diff2 = replace(diff2, mincov2 < 20, NA)) %>%
  mutate(diff3 = replace(diff3, mincov3 < 20, NA)) -> mismatchdiff
  
# reduce this tibble down to the bare minimum
mismatchdiff %>%
  filter(strand == "+") %>%
  ungroup() %>%
  select(anticodon, adjpos, diff1, diff2, diff3) %>%
  filter(adjpos > 0 ) -> small # just drop the 5´-adapter data. harder for 3´ adapter b/c length differences

```

Plotting code block. If the tRNAs all had the same length I'd write a plotting function, but they don't. There's just enough tweaking by hand of tick marks, etc. for the final figures to leave this code as-is. To plot different tRNAs, filter by the appropraite anticodon in the code block below and adjust the plotting window as needed. tRNA lengths (which include 5' and 3' adapters) can be pulled from the alignment reference: /ref/noints-consensus-ONTadapt-sacCer3-tRNAs.fa
```{r}
# select a tRNA 
small %>%
  filter(anticodon == "GCT") %>%
  select(-anticodon) -> smol

# reshape::melt this into a long, tidy puddle of data
# melt does not like being piped into? so we do this separately
melt(smol, id = c("adjpos")) -> longsmol

# plot stuff. remember to change the title (and the y limits / ticks if needed)!
## and tRNAs have Us not Ts, so we change the names accordingly
ggplot(longsmol, aes(x = variable,
                     y = adjpos,
                     fill = value)) +
  geom_tile(color = "black",
            lwd = .15,
            linetype = 1)+
  scale_fill_gradient2(low = "#0072B2",
                       mid = "#FFFFFF",
                       high = "#D55E00",
                       na.value = 'gray') +
  coord_fixed() +
  theme(plot.background=element_blank(),
        panel.grid = element_blank(),
        panel.border=element_blank(),
        panel.background = element_blank(),
        axis.text.x = element_blank(),
  #      axis.text.x = element_text(angle = 90),
        axis.ticks.x = element_blank(),
        legend.position = "left",
        legend.direction="vertical",
        legend.margin=margin(grid::unit(0, "cm")),
        legend.key.height=grid::unit(0.4, "cm"),
        legend.key.width=grid::unit(0.2, "cm")) +
  labs(title = "Ser-GCU",
       x = "",
       y = "",
       fill = "∆MM") +
  scale_y_reverse(
#    limits = c(84,0),
    breaks = c(90, 85, 80, 75, 70, 65, 60, 55, 50, 45, 40, 35, 30, 25, 20, 15, 10, 5)
  )  + 
  scale_x_discrete(expand=c(0,0))+
  coord_cartesian(xlim = c(0,4)) +
  coord_fixed(ratio = 1/2) 


```

