---
title: "Nanopolish eventalign parsing"
output: html_document
date: '2022-04-20'
---

```{r}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(cowplot)
library(colorblindr)
library(here)
here::i_am("rmd/Fig4EFGH_eventalign_parse.Rmd")
```


```{r}
read_delim(here("data/download_zenodo/oligos_eventalign/ctrl.bwa.eventalign_output.txt"), delim = "\t", col_names = T) %>%
  mutate(id = "unmodified") %>%
  select(-strand, -contig, -model_kmer, -model_mean, -model_stdv, -standardized_level) -> ctrlsplint

read_delim(here("data/download_zenodo/oligos_eventalign/2P.bwa.eventalign_output.txt"), delim = "\t", col_names = T) %>%
  mutate(id = "modified")  %>%
  select(-strand, -contig, -model_kmer, -model_mean, -model_stdv, -standardized_level) -> modsplint

# add 3 nt to account for nanopolish working in kmer space & beginning count w/ 0
bind_rows(ctrlsplint, modsplint) %>%
  mutate(position = position +3) -> splint_np
```

Now take that tibble and split it out into a few different ones, because we want to do several distinct things for each read:
* Sum the event_length by position (collapsing positions with multiple events) and multiply by 1000 to get Dwell in milliseconds
* Take the mean of event_level by position (collapsing positions with multiple events)

```{r}
splint_np %>%
  group_by(position, reference_kmer, read_index, id) %>%
  summarize(mean_current = mean(event_level_mean)) -> splint_current
```
I should be able to make the exact same plot as before.
```{r}
splint_current %>%
  filter(position > 29 & position < 55) %>%
  ggplot(., aes(x = position, 
                y = mean_current, 
                group = desc(read_index),
                color = id)) +
  geom_line(aes(color = id), alpha=.05) +
  scale_color_manual(values = c(unmodified = "#56B4E9", modified = "#E69F00")) +
  theme_minimal_hgrid() +
  scale_x_continuous(breaks = seq(0, 85, by = 5), expand = c(0, .5)) +
  theme(legend.position = c(.01, 1),
        legend.box.background = element_rect(color = "transparent", fill='white'),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  geom_vline(xintercept=41, linetype="dashed", color = "black", size=.5) +
  geom_vline(xintercept=34, linetype="dashed", color = "gray", size=.5) +
  geom_vline(xintercept=48, linetype="dashed", color = "gray", size=.5) +
  theme(aspect.ratio=1/4)+
  labs(title = "",
       x = "",
       y = "Current intensity (pA)",
       fill = "",
       color = "")
ggsave("perread_CI_on_splint.png")
```
Ok, yep, that's the same. So now we can do a similar thing for Dwell.
```{r}
splint_np %>%
  group_by(position, reference_kmer, read_index, id) %>%
  summarize(Dwell = sum(event_length)*1000) -> splint_dwell

splint_dwell %>%
  ggplot(., aes(x = position, 
                y = Dwell, 
                group = desc(read_index),
                color = id)) +
  geom_line(aes(color = id), alpha=.2) +
  scale_color_manual(values = c(unmodified = "#56B4E9", modified = "#E69F00")) +
  theme_minimal_hgrid() +
  scale_x_continuous(breaks = seq(0, 85, by = 5), expand = c(0, .5)) +
  theme(legend.position = c(.01, 1),
        legend.box.background = element_rect(color = "transparent", fill='white'),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  theme(aspect.ratio=1/4)+
  geom_vline(xintercept=41, linetype="dashed", color = "black", size=.5) +
  geom_vline(xintercept=34, linetype="dashed", color = "gray", size=.5) +
  geom_vline(xintercept=48, linetype="dashed", color = "gray", size=.5) +
  labs(title = "",
       x = "",
       y = "Dwell (ms)",
       fill = "",
       color = "")
```
It would also be nice to see mean dwell, so let's do that.
```{r}
splint_dwell %>%
  group_by(position, reference_kmer, id) %>%
  summarize(Dwell = mean(Dwell)) -> mean_dwell


mean_dwell %>%
 # filter(position > 29 & position < 56) %>%
  ggplot(., aes(x = position, 
                y = Dwell, 
                color = id)) +
  geom_line(aes(color = id)) +
  scale_color_manual(values = c(unmodified = "#56B4E9", modified = "#E69F00")) +
  theme_minimal_hgrid() +
  scale_x_continuous(breaks = seq(0, 85, by = 5), expand = c(0, .5)) +
  theme(legend.position = c(.01, 1),
        legend.box.background = element_rect(color = "transparent", fill='white'),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  theme(aspect.ratio=1/4)+
  geom_vline(xintercept=41, linetype="dashed", color = "black", size=.5) +
  geom_vline(xintercept=52, linetype="dashed", color = "#D55E00", size=.5) +
  labs(title = "",
       x = "",
       y = "Mean Dwell (ms)",
       fill = "",
       color = "")
```
Jay made a good suggestion, reworking the per-read Dwell into a box plot like I did for Q scores.
```{r}
splint_dwell %>%
  filter(position > 34 & position < 61) %>%
  group_by(id) %>%
  ggplot(., aes(x = as.factor(position), y = Dwell, fill = fct_rev(id))) +
  geom_boxplot(outlier.shape = NA) +
  scale_x_discrete(breaks = seq(0, 85, by = 5), expand = c(0, 0)) +
  scale_y_log10() +
  scale_fill_manual(values = c("#56B4E9", "#E69F00"), limits = c("unmodified", "modified")) +
  theme_minimal_hgrid() +
  theme(legend.position = c(.05, .9),
        legend.box.background = element_rect(color = "transparent", fill='white'),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  theme(aspect.ratio=1/3) +
  geom_vline(xintercept=7, linetype="dashed", color = "black", size=1, alpha = .5) + # modified position
  labs(title = "",
       x = "",
       y = "Dwell (ms)",
       fill = "")
ggsave("splint_perread_dwell.png")
```
Now to do the same thing for HAC1.
```{r}
read_delim(here("data/download_zenodo/mRNA_eventalign/YFL031W_mRNA.CRY1_Tm_NoExo.eventalign_output.txt"), delim = "\t", col_names = c("contig", "position",	"reference_kmer",	"read_id",	"strand",	"event_index",	"event_level_mean",	"event_stdv",	"event_length",	"model_kmer",	"model_mean",	"model_stdv",	"standardized_level")) %>%
  mutate(id = "wild-type") %>%
  select(-strand, -contig, -model_kmer, -model_mean, -model_stdv, -standardized_level) -> hac1_wildtype

read_delim(here("data/download_zenodo/mRNA_eventalign/YFL031W_mRNA.xrn1tpt1_Tm_NoExo.eventalign_output.txt"),, delim = "\t", col_names = c("contig", "position",	"reference_kmer",	"read_id",	"strand",	"event_index",	"event_level_mean",	"event_stdv",	"event_length",	"model_kmer",	"model_mean",	"model_stdv",	"standardized_level")) %>%
  mutate(id = "mutant")  %>%
  select(-strand, -contig, -model_kmer, -model_mean, -model_stdv, -standardized_level) -> hac1_mutant

bind_rows(hac1_wildtype, hac1_mutant) %>%
  mutate(position = position + 3) -> hac1_np

hac1_np %>%
  group_by(position, reference_kmer, read_id, id) %>%
  summarize(Dwell = sum(event_length)*1000) -> hac1_dwell

hac1_dwell %>%
  filter(position > 719 & position < 746) %>%
  group_by(id) %>%
  ggplot(., aes(x = as.factor(position), y = Dwell, fill = id)) +
  geom_boxplot(outlier.shape = NA) +
  scale_x_discrete(breaks = seq(690, 750, 5), expand = c(0,0)) +
  scale_y_log10() +
  scale_fill_manual(values = c("#56B4E9", "#E69F00"), limits = c("wild-type", "mutant")) +
  theme_minimal_hgrid() +
  theme(legend.position = c(.05, .9),
        legend.box.background = element_rect(color = "transparent", fill='white'),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  theme(aspect.ratio=1/3) +
  geom_vline(xintercept=8, linetype="dashed", color = "black", size=1, alpha = .5) +
  labs(title = "",
       x = "",
       y = "Dwell (ms)",
       fill = "")

ggsave("hac1_perread_dwell.png")
```
Here is mean dwell on HAC1, also.
```{r}
hac1_np %>%
  group_by(position, reference_kmer, read_id, id) %>%
  summarize(Dwell = sum(event_length)*1000) %>%
  group_by(position, reference_kmer, id) %>%
  summarize(Dwell = mean(Dwell)) -> mean_dwell

mean_dwell %>%
  ggplot(., aes(x = position, 
                y = Dwell, 
                color = id)) +
  geom_line(aes(color = id)) +
  scale_color_manual(values = c('wild-type' = "#56B4E9", 'mutant' = "#E69F00")) +
  theme_minimal_hgrid() +
  scale_x_continuous(breaks = seq(0, 1200, by = 100), expand = c(0, .5)) +
  theme(legend.position = c(.01, 1),
        legend.box.background = element_rect(color = "transparent", fill='white'),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  theme(aspect.ratio=1/4)+
  labs(title = "",
       x = "",
       y = "Mean Dwell (ms)",
       fill = "",
       color = "")
ggsave("HAC1_meandwell.png")
```
Then replicate the ∆ current intensity plot with a ∆ dwell plot 
```{r}
# AGAIN NEED TO COMPENSATE FOR NANOPOLISH POSITION EFFECT BY ADDING +3
hac1_wildtype %>%
  mutate(position = position +3) %>% 
  select(position, read_id, event_length, id) %>%
  group_by(position, read_id, id) %>%
  summarize(Dwell = sum(event_length)*1000) %>%
  ungroup() %>%
  group_by(position, id) %>%
  summarize(Dwell = mean(Dwell)) -> wt_collapsed

hac1_mutant %>%
  mutate(position = position +3) %>% 
  select(position, read_id, event_length, id) %>%
  group_by(position, read_id, id) %>%
  summarize(Dwell = sum(event_length)*1000) %>%
  ungroup() %>%
  group_by(position, id) %>%
  summarize(Dwell = mean(Dwell)) -> mt_collapsed

full_join(wt_collapsed, mt_collapsed, by = "position") %>%
  mutate(delta = Dwell.x - Dwell.y) -> nanopolish_collapsed

nanopolish_collapsed %>%
  filter(position > 719 & position < 746) %>%
  ggplot(., aes(x = position, y = delta)) +
  geom_col(show.legend = F, fill = "#009E73") +
  scale_x_continuous(breaks = seq(0, 1200, 5), expand = c(0, .5)) +
  theme_minimal_hgrid() +
  theme(aspect.ratio=1/3)+
  labs(title = "",
       x = "",
       y = "∆ dwell (ms) \n (WT-MT)")
ggsave("hac1_delta_dwell.png")
```

